The promise of community-driven preprints in ecology and evolution

### 2024-06-20 ms/ms.qmd
200 words

# Abstract
Publishing preprints has become an entrenched practice across a multitude of scientific disciplines and it is quickly becoming commonplace in ecology and evolutionary biology. Preprints can facilitate the rapid sharing of scientific knowledge to establish precedence, speed up the dissemination of research findings, and enable feedback from the research community before peer review. Yet, significant barriers to preprint use exist including language barriers, a lack of understanding about the benefits of preprints and a lack of diversity in the types of research outputs accepted (e.g., reports). Community driven preprint initiatives can allow a research community to come together to break down these barriers and move scientific publishing practices into new and exciting directions that promote greater equity and better coverage of global knowledge.  Here, we explore the first preprints uploaded to *EcoEvoRxiv*, a community-driven preprint server for ecologists and evolutionary biologists, to characterise preprint practices in ecology, evolution and conservation. Our perspective piece highlights some of the unique initiatives that *EcoEvoRxiv* has taken to break down barriers to scientific publishing by exploring the composition of articles, how gender and career stage influence preprint use, whether preprints are associated with greater open science principles (e.g., code and data sharing), and tracking preprint publication outcomes. Our analysis identifies areas that we still need to improve upon but highlight how community-driven initiatives, such as *EcoEvoRxiv*, can play a crucial role in shaping publishing practices in biology.

>>>>

# Abstract
Publishing preprints is quickly becoming commonplace in ecology and evolutionary biology. Preprints can facilitate the rapid sharing of scientific knowledge establishing precedence and enabling feedback from the research community before peer review. Yet, significant barriers to preprint use exist including language barriers, a lack of understanding about the benefits of preprints and a lack of diversity in the types of research outputs accepted (e.g., reports). Community driven preprint initiatives can allow a research community to come together to break down these barriers to improve equity and coverage of global knowledge. Here, we explore the first preprints uploaded to EcoEvoRxiv, a community-driven preprint server for ecologists and evolutionary biologists, to characterise preprint practices in ecology, evolution and conservation. Our perspective piece highlights some of the unique initiatives that EcoEvoRxiv has taken to break down barriers to scientific publishing by exploring the composition of articles, how gender and career stage influence preprint use, whether preprints are associated with greater open science principles (e.g., code and data sharing), and tracking preprint publication outcomes. Our analysis identifies areas that we still need to improve upon but highlight how community-driven initiatives, such as EcoEvoRxiv, can play a crucial role in shaping publishing practices in biology.
